### 2021-04-28 citation_fetching.Rmd
fetching citation info

---
title: "Fetch Citation Info"
author: "Mark Ziemann"
date: "28/04/2021"
output: html_document
---

## Intro

Here we are going to fetch the number of citations as well as the publication date for every PMCID.

The code below fetches the number of times that A PMC has been cited.

The output format is

1. PMCID

2. PMID

3. Number of cites

```{r,reutils2}

get_cites <- function(PMCID) {
  mylink <- elink(uid=PMCID , dbFrom="pmc", dbTo = "pubmed")
  PMID <- mylink[1]
  CITES <- elink(uid = PMID , dbFrom = "pubmed"  , linkname = "pubmed_pubmed_citedin" )
  N_CITES <- length(CITES[1])
  return(c(PMCID,PMID,N_CITES))
}

PMCID=3155154

get_cites(PMCID = PMCID)

```