### 2021-04-29 citation_fetching.Rmd
script works

---
title: "Fetch PMC Citation Info"
author: "Mark Ziemann"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    fig_width: 7
    fig_height: 7
theme: cosmo
---

## Intro

Here we are going to fetch the number of citations as well as the publication date for every PMCID.

The code below fetches the number of times that A PMC has been cited.

The output format is

1. PMCID

2. PMID

3. Number of cites

```{r,reutils2}

library("reutils")

get_cites <- function(PMCID) {
  mylink <- elink(uid=PMCID , dbFrom="pmc", dbTo = "pubmed")
  PMID <- mylink[1]
  CITES <- elink(uid = PMID , dbFrom = "pubmed"  , linkname = "pubmed_pubmed_citedin" )
  N_CITES <- length(CITES[1])
  return(c(PMCID,PMID,N_CITES))
}

PMCID=3155154

get_cites(PMCID = PMCID)

```

## Try the following

Now to get a bunch of these.

```{r,reutils3}

PMCIDS <- c("PMC6317205", "PMC6317219", "PMC6318854", "PMC6318897", "PMC6322516", "PMC6323737", "PMC6323741", "PMC6325784", "PMC6325795", "PMC6325889", "PMC6326702", "PMC6327033",
            "PMC6327913", "PMC6328174", "PMC6328465", "PMC6328499", "PMC6329100", "PMC6329100", "PMC6329125", "PMC6329125", "PMC6329193", "PMC6329348")

PMCIDS <- as.numeric(gsub("PMC","",PMCIDS))

res <- sapply(X = PMCIDS, FUN = get_cites)

res
```

Looks like it works :)

## Session information

```{r,sessioninfo}

sessionInfo()

```