### 2021-07-15 cleaning_aid.Rmd
started cleaning

---
title: "Cleaning Data"
author: "Mark Ziemann & Kaumadi Wijesooriya"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    fig_width: 7
    fig_height: 7
theme: cosmo
---

## Intro

here is some background info blah blah

```{r,load data}

x <- read.table("PMC 2019 - analysis.tsv",header=TRUE,fill=TRUE,sep="\t")
head(x)
colnames(x)

# omics type
omics <- x$Omics.type
omics_split <- strsplit(omics,", ")
omics <- unlist(omics_split)
table(omics)

# organism
org <- x$Organism
org_split <- strsplit(org,", ")
org <- unlist(org_split)
table(org)

```