### 2021-08-09 analysis.Rmd
minor improvements

---
title: "PMC Survey Data Analysis"
author: "Mark Ziemann & Kaumadi Wijesooriya"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    fig_width: 7
    fig_height: 5
theme: cosmo
---

Source: https://github.com/markziemann/SurveyEnrichmentMethods

## Intro

Here we are performing an analysis of the ways enrichment analysis has been done in a sample of 1500 PMC articles from 2019.

```{r,begin}

knitr::opts_chunk$set(fig.width=7, fig.height=5)

library("wordcloud")
library("RColorBrewer")
library("wordcloud2")
library("reutils")
library("XML")
library("kableExtra")
library("vioplot")

```

## Overview of included and excluded analyses

```{r,exclude1}

x <- read.table("PMC_2019-analysis.tsv",header=TRUE,fill=TRUE,sep="\t")
head(x)
colnames(x)
dim(x)

exclude <- subset(x,x$GS.version=="EXCLUDE")
nrow(exclude)
length(unique(exclude$Pubmed.Central.ID))

x <- subset(x,x$GS.version!="EXCLUDE")
nrow(x)
length(unique(x$Pubmed.Central.ID))

```

## Journal

```{r,Journal1}

journal <- x$Journal
journal_split <- strsplit(journal,", ")
journal <- unlist(journal_split)
res <- table(journal)
res <- res[order(res)]
length(res)
res

par(mar=c(1,1,1,1))
#names(res) <- gsub("Gene expression array","RNA array",names(res))
wordcloud(words = names(res), freq = res, min.freq = 1, 
    max.words=200, random.order=FALSE, rot.per=0.35, 
    colors=brewer.pal(8, "Dark2"))

par(mar=c(5,12,3,1))
barplot(tail(res,20),horiz=TRUE,las=1,cex.names = 0.7, xlab="no. analyses",
        main = "Journal", xlim=c(0,100))
grid()

other <- sum(res[1:(nrow(res)-10)])
res2 <- c(other,tail(res,9))
names(res2)[1] <- "Other"
par(mar=c(5,12,3,1))
barplot(res2,horiz=TRUE,las=1,cex.names = 0.7, xlab="no. analyses",
        main = "Journal", xlim=c(0,1200))
grid()

```

## Omics type

```{r,omics1}

omics <- x$Omics.type
omics_split <- strsplit(omics,", ")
omics <- unlist(omics_split)
res <- table(omics)
res <- res[order(res)]
length(res)
res

par(mar=c(1,1,1,1))
names(res) <- gsub("Gene expression array","RNA array",names(res))
wordcloud(words = names(res), freq = res, min.freq = 1, 
    max.words=200, random.order=FALSE, rot.per=0.35, 
    colors=brewer.pal(8, "Dark2"))

par(mar=c(5,12,3,1))
barplot(tail(res,20),horiz=TRUE,las=1,cex.names = 0.7, xlab="no. analyses",
        main = "Omics type", xlim=c(0,600))
grid()

other <- sum(res[1:(nrow(res)-10)])
res2 <- c(other,tail(res,9))
names(res2)[1] <- "Other"
par(mar=c(5,12,3,1))
barplot(res2,horiz=TRUE,las=1,cex.names = 0.7, xlab="no. analyses",
        main = "Omics type", xlim=c(0,600))
grid()

```

## Organism

```{r,org1}

org <- x$Organism
org_split <- strsplit(org,", ")
org <- unlist(org_split)
res <- table(org)
res <- res[order(res)]
length(res)
res

par(mar=c(1,1,1,1))
names(res) <- gsub("Homo sapiens","human",names(res))
wordcloud(words = names(res), freq = res, min.freq = 1, 
    max.words=200, random.order=FALSE, rot.per=0.35, 
    colors=brewer.pal(8, "Dark2"), scale=c(4,.5))

par(mar=c(5,12,3,1))
names(res) <- gsub("human","Homo sapiens",names(res))
barplot(tail(res,20),horiz=TRUE,las=1,cex.names = 0.7, xlab="no. analyses",
        main = "Organism", xlim=c(0,1200))
grid()

other <- sum(res[1:(nrow(res)-10)])
res2 <- c(other,tail(res,9))
names(res2)[1] <- "Other"
par(mar=c(5,12,3,1))
barplot(res2,horiz=TRUE,las=1,cex.names = 0.7, xlab="no. analyses",
        main = "Organism", xlim=c(0,1200))
grid()

```

## Gene set library

```{r,gsl1}

GSL <-x$Gene.set.library
GSL_split <- strsplit(GSL,", ")
GSL <- unlist(GSL_split)
res <- table(GSL)
res <- res[order(res)]
length(res)
which(names(res)=="Not stated")/sum(res)*100
res

par(mar=c(1,1,1,1))
names(res) <- gsub("Homo sapiens","human",names(res))
wordcloud(words = names(res), freq = res, min.freq = 1, 
    max.words=200, random.order=FALSE, rot.per=0.35, 
    colors=brewer.pal(8, "Dark2"))

par(mar=c(5,12,3,1))
barplot(tail(res,20),horiz=TRUE,las=1,cex.names = 0.7, xlab="no. analyses",
        main = "Gene set library", xlim=c(0,1000))
grid()

other <- sum(res[1:(nrow(res)-10)])
res2 <- c(other,tail(res,9))
names(res2)[1] <- "Other"
par(mar=c(5,12,3,1))
barplot(res2,horiz=TRUE,las=1,cex.names = 0.7, xlab="no. analyses",
        main = "Gene set library", xlim=c(0,500))
grid()

```

## Gene set version

```{r,gsv1}

GSV <-x$GS.version
res <- table(GSV)
res
res[1]/sum(res)*100

par(mar=c(5,12,3,1))
barplot(tail(res,20),horiz=TRUE,las=1,cex.names = 1, xlab="no. analyses",
        main = "Gene set version defined", xlim=c(0,1600))
grid()

```

## Statistical test used

```{r,test1}

test <-x$Statistical.test.used
test <- strsplit(test,", ")
test <- unlist(test)
res <- table(test)
res <- res[order(res)]
res[which(names(res)=="Not stated")] / sum(res) * 100
res

par(mar=c(1,1,1,1))
wordcloud(words = names(res), freq = res, min.freq = 1, 
    max.words=200, random.order=FALSE, rot.per=0.35, 
    colors=brewer.pal(8, "Dark2"))

par(mar=c(5,12,3,1))
barplot(tail(res,20),horiz=TRUE,las=1,cex.names = 0.7, xlab="no. analyses",
        main = "Test used", xlim=c(0,1200))
grid()

other <- sum(res[1:(nrow(res)-10)])
res2 <- c(other,tail(res,9))
names(res2)[1] <- "Other"
par(mar=c(5,12,3,1))
barplot(res2,horiz=TRUE,las=1,cex.names = 0.7, xlab="no. analyses",
        main = "Test used", xlim=c(0,1200))
grid()

```

## FDR Correction

```{r,fdr1}

fdr <-x$FDR.Correction
fdr <- strsplit(fdr,", ")
fdr <- unlist(fdr)
res <- table(fdr)
res <- res[order(res)]
sum(res[which(names(res)!="Yes")])/sum(res)*100
res

par(mar=c(1,1,1,1))
wordcloud(words = names(res), freq = res, min.freq = 1, 
    max.words=200, random.order=FALSE, rot.per=0.35, 
    colors=brewer.pal(8, "Dark2"))

par(mar=c(5,12,3,1))
barplot(tail(res,20),horiz=TRUE,las=1,cex.names = 0.7, xlab="no. analyses",
        main = "FDR", xlim=c(0,800))
grid()

```

## App used

```{r,app1}

App <-x$App.used
App_split <- strsplit(App,", ")
App <- unlist(App_split)
res <- table(App)
res <- res[order(res)]
res[which(names(res)=="Not stated")]/sum(res)*100
res

par(mar=c(1,1,1,1))
wordcloud(words = names(res), freq = res, min.freq = 1, 
    max.words=200, random.order=FALSE, rot.per=0.35, 
    colors=brewer.pal(8, "Dark2"))

par(mar=c(5,12,3,1))
barplot(tail(res,20),horiz=TRUE,las=1,cex.names = 0.7, xlab="no. analyses",
        main = "App used", xlim=c(0,400))
grid()

other <- sum(res[1:(nrow(res)-10)])
res2 <- c(other,tail(res,9))
names(res2)[1] <- "Other"
par(mar=c(5,12,3,1))
barplot(res2,horiz=TRUE,las=1,cex.names = 0.7, xlab="no. analyses",
        main = "App used", xlim=c(0,500))
grid()

```

## App version

```{r,appv1}

APV <-x$App.Version
res <- table(APV)
res
res[1]/sum(res)*100

par(mar=c(5,12,3,1))
barplot(tail(res,20),horiz=TRUE,las=1,cex.names = 1, xlab="no. analyses",
        main = "App version defined", xlim=c(0,1600))
grid()

```

## Code available

```{r,code1}

code <-x$Code.availability
res <- table(code)
res
res[1]/sum(res)*100

par(mar=c(5,12,3,1))
barplot(tail(res,20),horiz=TRUE,las=1,cex.names = 1, xlab="no. analyses",
        main = "Code availability", xlim=c(0,300))
grid()

```

## Background gene set

```{r,background1}

BG <-x$Background.gene.set
res <- table(BG)
res
sum(res[which(names(res)!="Yes")])/sum(res)*100

par(mar=c(5,12,3,1))
barplot(tail(res,20),horiz=TRUE,las=1,cex.names = 1, xlab="no. analyses",
        main = "Background list defined", xlim=c(0,1200))
grid()

```

## Gene lists provided

```{r,genelists1}

GL <-x$Gene.lists.provided
res <- table(GL)
res
sum(res[which(names(res)!="Yes")])/sum(res)*100

par(mar=c(5,12,3,1))
barplot(tail(res,20),horiz=TRUE,las=1,cex.names = 1, xlab="no. analyses",
        main = "Gene lists provided", xlim=c(0,1200))
grid()
colnames(x)

```

## Assumptions violated

```{r,assumptions1}

ok <- nrow(subset(x,Assumptions.violated=="No"))
ok
bad <- nrow(subset(x,Assumptions.violated!="No"))
bad
ok/sum(bad,ok)*100

ass <-x$Assumptions.violated
ass <- strsplit(ass,", ")
ass <- unlist(ass)
res <- table(ass)
res <- res[order(res)]
res

par(mar=c(1,1,1,1))
wordcloud(words = names(res), freq = res, min.freq = 1, 
    max.words=200, random.order=FALSE, rot.per=0.35, 
    colors=brewer.pal(8, "Dark2"))

par(mar=c(5,12,3,1))
barplot(tail(res,20),horiz=TRUE,las=1,cex.names = 1, xlab="no. analyses",
        main = "Assumptions violated", xlim=c(0,1400))
grid()

```

## Score

Here I am proposing a scoring scheme.
If there is information missing from the article, or basic mistakes are made, then a point is deducted.
If the article goes over and above the basic reproducibility, then they are awarded a point.

Gene set library origin not stated = -1

Gene set library version not stated = -1

Stat test not stated = -1

No stat test conducted = -1

No FDR correction conducted = -1

App used not stated = -1

App version not stated = -1

Background list not defined = -1

Inappropriate background list used = -1

Code availability = +1

Gene lists provided = +1

```{r,score}

score <- function(r){
  r[is.na(r)] <- 0
  SCORE=0
  # gene set lib
  if(r[7]=="Not stated"){
    SCORE=SCORE-1
  }
  # GS version
  if(r[8]=="No"){
    SCORE=SCORE-1
  }
  # stat test
  if(r[9]=="No test"){
    SCORE=SCORE-1
  }
  if(r[9]=="Not stated"){
    SCORE=SCORE-1
  }
  # FDR
  if(r[10]!="Yes"){
    SCORE=SCORE-1
  }
  # app used
  if(r[11]=="Not stated"){
    SCORE=SCORE-1
  }
  # App version
  if(r[12]=="No"){
    SCORE=SCORE-1
  }
  # Code availability
  if(r[13]=="Yes"){
    SCORE=SCORE+1
  }
  # Background
  if(r[14]=="Not stated"){
    SCORE=SCORE-1
  }
  if(r[14]=="No"){
    SCORE=SCORE-1
  }
  if(r[14]=="Stated, but incorrect"){
    SCORE=SCORE-1
  }
  # gene list
  if(r[16]=="Yes"){
    SCORE=SCORE+1
  }
  return(SCORE)
}

scores <- apply(X = x,MARGIN = 1, FUN = score)
hist(scores)
x$scores <- scores

```

## Journal rank

```{r,journal rank}

jmetrics <- read.table("scimagojr_2020.csv",sep=";",header=TRUE,dec = ",")
jmetrics$Title <- toupper(jmetrics$Title)
head(jmetrics,3)
dim(jmetrics)

nlmcat <- readLines("nlmcatalog_result.txt")
journaltitle <- nlmcat[grep("Title\\(s\\):",nlmcat)]
journaltitle <- gsub("Title\\(s\\): ","",journaltitle)
journaltitle <- gsub("\\.$","",journaltitle)
journaltitle <- toupper(journaltitle) 
journalabbrev <- nlmcat[grep("Title Abbreviation:",nlmcat)]
journalabbrev <- sapply(strsplit(journalabbrev,":"),"[[",2)
journalabbrev <- gsub(" $","",journalabbrev)
journalabbrev <- gsub("^ ","",journalabbrev)
jdf <- data.frame(journalabbrev,journaltitle)

mjdf <- merge(jdf,jmetrics,by.x="journaltitle",by.y="Title")

xm <- merge(x,mjdf,by.x="Journal",by.y="journalabbrev",all.x = TRUE)

```

## Analysis scores vs Journal rank

```{r,scoresvsjrank1}

tail(xm$Journal)

mylm1 <- lm (xm$scores ~ xm$SJR)
plot(xm$SJR,xm$scores,xlab="SJR",ylab="score",main="Association of analysis scores with journal metrics")
abline(mylm1,col="red")

cor.test(xm$scores,xm$SJR,method="pearson")

cor.test(xm$scores,xm$SJR,method="spearman")

```
## Best and worst journals

Minimum of 5 analyses.

```{r,goodbadugly}

xmm <-xm[,c(1,17)]
tab <- table(xmm$Journal)
tab <- tab[which(tab>5)]
xmm <- xmm[which(xmm$Journal %in% names(tab)),]
dim(xmm)
jres <- aggregate(. ~ Journal, xmm ,mean)
jsd <- aggregate(. ~ Journal, xmm ,sd)
jres$sd <- jsd$scores
jres <- jres[order(jres$scores),]
dim(jres)
head(jres)

par(mar=c(5,12,3,1))
barplot(tail(jres$scores,20),names.arg = tail(jres$Journal,20),horiz=TRUE,las=1,cex.names = 0.7, xlab="mean score",
        main = "Highest scoring journals",xlim=c(-4,0))
grid()

par(mar=c(5,12,3,1))
barplot(head(jres$scores,20),names.arg = head(jres$Journal,20),horiz=TRUE,las=1,cex.names = 0.7, xlab="mean score",
        main = "Lowest scoring journals",xlim=c(-5,0))
grid()

j <- unique(xmm$Journal)
jscores <- sapply(j,function(jj) {
 xmm[which(xmm$Journal == jj ),2]  
})

jscores <- jscores[order(unlist(lapply(jscores,mean)))]

par(mar=c(5,12,3,1))
vioplot(tail(jscores,20),horizontal = TRUE,las=1,cex.axis=0.75,main="Highest scoring journals")

par(mar=c(5,12,3,1))
vioplot(head(jscores,20),horizontal = TRUE,las=1,cex.axis=0.75,main="Lowest scoring journals")

```

## Mean analysis scores vs Journal rank

```{r,scoresvsjrank2}

jjres <- merge(jres,mjdf,by.x="Journal",by.y="journalabbrev",all.x = TRUE)

mylm2 <- lm (jjres$scores ~ jjres$SJR)
plot(jjres$SJR,jjres$scores,xlab="SJR",ylab="score",main="Association of mean analysis score with journal metrics")
abline(mylm2,col="red")

cor.test(jjres$scores,jjres$SJR,method="pearson")

cor.test(jjres$scores,jjres$SJR,method="spearman")

```

## Article citations

Get the number of citations per article.

```{r,cites}

if (file.exists("cites_data.rds") ) {
  cites <- readRDS("cites_data.rds")
} else {
  PMCIDS <- as.numeric(gsub("PMC","",x$Pubmed.Central.ID ))
  mylink <- elink(uid=PMCIDS , dbFrom="pmc", dbTo = "pubmed")
  PMID <- mylink[1]
  #writeLines(PMID,"pubmedids.txt")
  res <- sapply(X=PMID,FUN=function(x){
      Sys.sleep(1)
      CITES <- elink(uid = x , dbFrom = "pubmed"  , linkname = "pubmed_pubmed_citedin" ) 
      CITES=length(CITES[1])
      return(CITES)
   })
  mylink2 <- elink(uid=PMID , dbFrom="pubmed", dbTo = "pmc")
  cites <- data.frame(mylink2[1],PMID,res)
  colnames(cites) <- c("PMCID","PMID","cites")
  saveRDS(cites,file="cites_data.rds")
}

cites$PMCID <- gsub("^","PMC",cites$PMCID)
cites$PMID <- gsub("^","PMID",cites$PMID)
xm <- merge(x,cites,by.x="Pubmed.Central.ID",by.y="PMCID")
dim(xm)

```

Plot the number of citations against analysis scores

```{r,citesplot}

mylm2 <- lm (xm$scores ~ xm$cites)
plot(xm$cites,xm$scores,xlab="Num cites",ylab="score",main="Association of analysis scores with number of citations")
abline(mylm2,col="red")

cor.test(xm$scores , xm$cites, method="pearson")

cor.test(xm$scores , xm$cites, method="spearman")

mylm2 <- lm (xm$scores ~ log2(xm$cites))
plot(log2(xm$cites),xm$scores,xlab="log2(citations)",ylab="score",main="Association of analysis scores with log2(citations)")
abline(mylm2,col="red")

cor.test(xm$scores , xm$cites, method="pearson")

cor.test(xm$scores , xm$cites, method="spearman")
```

## Session information

```{r,session}

sessionInfo()

```