### 2021-08-13 example.Rmd
example RNA-seq ORA

---
title: "Example gene set analysis"
author: "Mark Ziemann & Kaumadi Wijesooriya"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    fig_width: 7
    fig_height: 5
theme: cosmo
---

Source: https://github.com/markziemann/SurveyEnrichmentMethods

## Intro

Here we are performing an analysis of some gene expression data to demonstrate the difference between ORA and FCS methods and to highlight the differences caused by improper background gene set use.

```{r,begin}

suppressPackageStartupMessages({
library("getDEE2") 
library("DESeq2")
library("clusterProfiler")
library("mitch")
library("kableExtra")
library("eulerr")
})

```

## Get expression data

I'm using some RNA-seq data looking at the effect of hyperglycemia on hepatocytes.

```{r,getdata,fig.width=7,fig.height=7}

mdat<-getDEE2Metadata("hsapiens")
samplesheet <- mdat[grep("SRP128998",mdat$SRP_accession),]
samplesheet<-samplesheet[order(samplesheet$SRR_accession),]
samplesheet$HG<-as.factor(c(1,1,1,1,1,1,0,0,0,0,0,0))
samplesheet$VPA<-as.factor(c(0,0,0,1,1,1,0,0,0,1,1,1))
s1 <- subset(samplesheet,VPA==0)

w<-getDEE2("hsapiens",samplesheet$SRR_accession,metadata=mdat,legacy = TRUE)
x<-Tx2Gene(w)
x<-x$Tx2Gene

# save the genetable for later
gt<-w$GeneInfo[,1,drop=FALSE]
gt$accession<-rownames(gt)
head(gt)

# counts 
x1<-x[,which(colnames(x) %in% s1$SRR_accession)]

# filter out lowly expressed genes
x1<-x1[which(rowSums(x1)/ncol(x1)>=(10)),]
nrow(x)
nrow(x1)

# MDS plot
plot(cmdscale(dist(t(x1))), xlab="Coordinate 1", ylab="Coordinate 2", pch=19, col=s1$HG, main="MDS")

#run DESeq2 for LG vs HG
y <- DESeqDataSetFromMatrix(countData = round(x1), colData = s1, design = ~ HG)
y <- DESeq(y)
LGvHG <- results(y)
LGvHG<-as.data.frame(LGvHG[order(LGvHG$pvalue),])
rownames(LGvHG)<-sapply(strsplit(rownames(LGvHG),"\\."),"[[",1)
head(LGvHG)

```

```{r,deplots,fig.width=7,fig.height=7}

maplot <- function(de,contrast_name) {
  sig <-subset(de, padj < 0.05 )
  up <-rownames(subset(de, padj < 0.05 & log2FoldChange > 0))
  dn <-rownames(subset(de, padj < 0.05 & log2FoldChange < 0))
  GENESUP <- length(up)
  GENESDN <- length(dn)
  DET=nrow(de)
  SUBHEADER = paste(GENESUP, "up, ", GENESDN, "down", DET, "detected")
  ns <-subset(de, padj > 0.05 )
  plot(log2(de$baseMean),de$log2FoldChange, 
       xlab="log2 basemean", ylab="log2 foldchange",
       pch=19, cex=0.5, col="dark gray",
       main=contrast_name, cex.main=0.7)
  points(log2(sig$baseMean),sig$log2FoldChange,
         pch=19, cex=0.5, col="red")
  mtext(SUBHEADER,cex = 0.7)
}

make_volcano <- function(de,name) {
    sig <- subset(de,padj<0.05)
    N_SIG=nrow(sig)
    N_UP=nrow(subset(sig,log2FoldChange>0))
    N_DN=nrow(subset(sig,log2FoldChange<0))
    DET=nrow(de)
    HEADER=paste(N_SIG,"@5%FDR,", N_UP, "up", N_DN, "dn", DET, "detected")
    plot(de$log2FoldChange,-log10(de$padj),cex=0.5,pch=19,col="darkgray",
        main=name, xlab="log2 FC", ylab="-log10 pval", xlim=c(-6,6))
    mtext(HEADER)
    grid()
    points(sig$log2FoldChange,-log10(sig$padj),cex=0.5,pch=19,col="red")
}

maplot(LGvHG,"LG v HG")
make_volcano(LGvHG,"LG v HG")

```

## Gene sets from Reactome

```{r,reactome}

if (! file.exists("ReactomePathways.gmt")) {
  download.file("https://reactome.org/download/current/ReactomePathways.gmt.zip", 
    destfile="ReactomePathways.gmt.zip")
  unzip("ReactomePathways.gmt.zip")
}
genesets<-gmt_import("ReactomePathways.gmt")

```

## FCS with Mitch

```{r,mitch1}

m <- mitch_import(LGvHG,DEtype = "DEseq2", geneTable = gt)
head(m)
mres <- mitch_calc(m,genesets = genesets)
head(mres$enrichment_result) 
m_up <- subset(mres$enrichment_result,p.adjustANOVA<0.05 & s.dist > 0)[,1]
m_dn <- subset(mres$enrichment_result,p.adjustANOVA<0.05 & s.dist < 0)[,1]
length(m_up)
length(m_dn)

```

## ORA with clusterprofiler

```{r,cp1}

genesets2 <- read.gmt("ReactomePathways.gmt")
head(genesets2)

LGvHG_up <- rownames(subset(LGvHG,log2FoldChange>0,padj<0.05))
LGvHG_up <- unique(gt[which(rownames(gt) %in% LGvHG_up),1])
head(LGvHG_up)

LGvHG_dn <- rownames(subset(LGvHG,log2FoldChange<0,padj<0.05))
LGvHG_dn <- unique(gt[which(rownames(gt) %in% LGvHG_dn),1])
head(LGvHG_dn)

LGvHG_bg <- rownames(LGvHG)
LGvHG_bg <- unique(gt[which(rownames(gt) %in% LGvHG_bg),1])
length(LGvHG_bg)
head(LGvHG_bg)

c_up <- enricher(gene = LGvHG_up, universe = LGvHG_bg,  maxGSSize = 5000, TERM2GENE = genesets2)
c_up <- head(c_up,1000)
head(c_up)
c_up <- rownames(subset(c_up, p.adjust < 0.05))

c_dn <- enricher(gene = LGvHG_dn, universe = LGvHG_bg,  maxGSSize = 5000, TERM2GENE = genesets2)
c_dn <- head(c_dn,1000)
head(c_dn)
c_dn <- rownames(subset(c_dn, p.adjust < 0.05))

```

## ORA with clusterprofiler combining up and down

```{r,cp2}

LGvHG_de <- rownames(subset(LGvHG,padj<0.05))
LGvHG_de <- unique(gt[which(rownames(gt) %in% LGvHG_de),1])
head(LGvHG_de)

d_de <- enricher(gene = LGvHG_de, universe = LGvHG_bg,  maxGSSize = 5000, TERM2GENE = genesets2)
d_de <- head(d_de,1000)
d_de <- rownames(subset(d_de, p.adjust < 0.05))
head(d_de)

```

## ORA with clusterprofiler with whole genome background list

```{r,cp3}

LGvHG_bg <- w$GeneInfo$GeneSymbol
length(LGvHG_bg)
head(LGvHG_bg)

f_up <- enricher(gene = LGvHG_up, universe = LGvHG_bg,  maxGSSize = 5000, TERM2GENE = genesets2)
f_up <- head(f_up,1000)
head(f_up)
f_up <- rownames(subset(f_up, p.adjust < 0.05))

f_dn <- enricher(gene = LGvHG_dn, universe = LGvHG_bg, maxGSSize = 5000, TERM2GENE = genesets2)
f_dn <- head(f_dn,1000)
head(f_dn)
f_dn <- rownames(subset(f_dn, p.adjust < 0.05))

```

## ORA combined with clusterprofiler with whole genome background list

```{r,cp4}

e_de <- enricher(gene = LGvHG_de, universe = LGvHG_bg, maxGSSize = 5000, TERM2GENE = genesets2)
e_de <- head(e_de,1000)
head(e_de)
e_de <- rownames(subset(e_de, p.adjust < 0.05))

```

## Venn diagram comparison

```{r,venn1}

par(cex.main=0.5)

par(mar=c(2,2,2,2))

v0 <- list("ORA up"=c_up,"ORA dn"=c_dn,
           "ORA comb" = d_de)

plot(euler(v0),quantities = TRUE,main="effect of combining up and down regulated genes")

v1 <- list("FCS up"=m_up, "FCS dn"=m_dn,
           "ORA up"=c_up,"ORA dn"=c_dn)

plot(euler(v1),quantities = TRUE, main="FCS compared to ORA")

v2 <- list("ORA up"=c_up,"ORA dn"=c_dn, 
           "ORA* up"=f_up,"ORA* dn"=f_dn )

plot(euler(v2),quantities = TRUE, main="Effect of inappropriate background* (whole genome)")

vx <- list("ORA up"=c_up,"ORA dn"=c_dn,
           "ORA comb" = d_de, "ORA* comb" = e_de)

plot(euler(vx),quantities = TRUE, main="combining up and down genes and whole genome bg*")

v3 <- list("ORA up"=c_up,"ORA dn"=c_dn, 
           "ORA* up"=f_up,"ORA* dn"=f_dn ,
           "FCS up"=m_up, "FCS dn"=m_dn)

plot(euler(v3),quantities = TRUE, main="all sets")

png("images/fcs_ora1.png")
plot(euler(v1),quantities = TRUE, main="FCS vs ORA")
dev.off()

png("images/orabg1.png")
plot(euler(v2),quantities = TRUE, main="Effect of inappropriate background* (whole genome)")
dev.off()

png("images/oracomb1.png")
plot(euler(vx),quantities = TRUE, main="combining up and down genes and whole genome bg*")
dev.off()

```

## Jaccard calculation

```{r,jacc1}

# ORA vs ORA combined
dc <- length(intersect(d_de, c(c_up,c_dn))) / length(union(d_de, c(c_up,c_dn)))
dc

# ORA vs ORA* combined
ec <- length(intersect(e_de, c(c_up,c_dn))) / length(union(e_de, c(c_up,c_dn)))
ec

# FCS vs ORA
cm <- length(intersect(c(c_up,c_dn), c(m_up,m_dn))) / length(union(c(c_up,c_dn), c(m_up,m_dn)))
cm

m_up <- gsub("^","up ",m_up)
m_dn <- gsub("^","dn ",m_dn)
m_de <- union(m_up,m_dn)

c_up <- gsub("^","up ",c_up)
c_dn <- gsub("^","dn ",c_dn)
c_de <- union(c_up,c_dn)

f_up <- gsub("^","up ",f_up)
f_dn <- gsub("^","dn ",f_dn)
f_de <- union(f_up,f_dn)

# ORA vs ORA*
cf <- length(intersect(c_de, f_de )) / length(union(c_de, f_de))
cf

# FCS vs ORA*
mf <- length(intersect(m_de, f_de )) / length(union(m_de, f_de))
mf

dat <- c("FCS vs ORA"=cm,"ORA vs ORA*"=cf,"FCS vs ORA*"=mf, "ORA vs ORA comb"=dc, "ORA vs ORA* comb"=ec)

dat

barplot(dat,ylab="jaccard metric")

```

## Session information

```{r,session}

sessionInfo()

```