### 2021-08-30 QC_analysis2.Rmd
major manuscript changes

---
title: "PMC Survey QC Analysis"
author: "Mark Ziemann & Kaumadi Wijesooriya"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    fig_width: 7
    fig_height: 5
theme: cosmo
---

## Intro

Here we are performing an analysis of 200 articles which is randomly selected from 1500 PMC articles.

```{r,begin}

knitr::opts_chunk$set(fig.width=7, fig.height=5)

library("wordcloud")
library("RColorBrewer")
library("wordcloud2")
library("reutils")
library("XML")
library("kableExtra")
library("Biobase")
library("vioplot")

```

## Overview of included and excluded analyses

```{r,exclude1}

x <- read.table("data/QC-analysis2.tsv",header=TRUE,fill=TRUE,sep="\t")
head(x)
colnames(x)
dim(x)

exclude <- subset(x,x$GS.version=="EXCLUDE")
nrow(exclude)
length(unique(exclude$Pubmed.Central.ID))

x <- subset(x,x$GS.version!="EXCLUDE")
nrow(x)
length(unique(x$Pubmed.Central.ID))

```
## Journal

```{r,Journal1}

journal <- x$Journal
journal_split <- strsplit(journal,", ")
journal <- unlist(journal_split)
res <- table(journal)
res <- res[order(res)]
length(res)
res

par(mar=c(1,1,1,1))
#names(res) <- gsub("Gene expression array","RNA array",names(res))
wordcloud(words = names(res), freq = res, min.freq = 1, 
          max.words=200, random.order=FALSE, rot.per=0.35, 
          colors=brewer.pal(8, "Dark2"))

par(mar=c(5,12,3,5))
barplot(tail(res,20),horiz=TRUE,las=1,cex.names = 0.7, xlab="no. analyses",
        main = "Journal", xlim=c(0,17))
grid()

other <- sum(res[1:(nrow(res)-10)])
res2 <- c(other,tail(res,9))
names(res2)[1] <- "Other"
par(mar=c(5,12,3,5))
barplot(res2,horiz=TRUE,las=1,cex.names = 1, xlab="no. analyses",
        main = "Journal", xlim=c(0,165))
text(y = (1:length(res2)*1.2)-1.2  , x = res2 + 7, label = res2, pos = 3, cex = 1, col = "black")

dir.create("images")
png("images/journals1.png",width=400,height=300)
par(mar=c(5,12,3,3))
barplot(res2,horiz=TRUE,las=1,cex.names = 1, xlab="no. analyses",
        main = "Journal", xlim=c(0,170))
text(y = (1:length(res2)*1.2)-1.2  , x = res2 + 10, label = res2, pos = 3, cex = 1, col = "black")
dev.off()
```

## Omics type

```{r,omics1}

omics <- x$Omics.type
omics_split <- strsplit(omics,", ")
omics <- unlist(omics_split)
res <- table(omics)
res <- res[order(res)]
length(res)
res

par(mar=c(1,1,1,1))
names(res) <- gsub("Gene expression array","RNA array",names(res))
wordcloud(words = names(res), freq = res, min.freq = 1, 
          max.words=200, random.order=FALSE, rot.per=0.35, 
          colors=brewer.pal(8, "Dark2"))

par(mar=c(5,12,3,5))
barplot(tail(res,20),horiz=TRUE,las=1,cex.names = 0.7, xlab="no. analyses",
        main = "Omics type", xlim=c(0,100))
grid()

other <- sum(res[1:(nrow(res)-10)])
res2 <- c(other,tail(res,9))
names(res2)[1] <- "Other"
par(mar=c(5,12,3,5))
barplot(res2,horiz=TRUE,las=1,cex.names = 1, xlab="no. analyses",
        main = "Omics type", xlim=c(0,100))
text(y = (1:length(res2)*1.2)-1.2  , x = res2+5, label = res2, pos = 3, cex = 1, col = "black")

png("images/omics1.png",width=400,height=300)
par(mar=c(5,12,3,3))
barplot(res2,horiz=TRUE,las=1,cex.names = 1, xlab="no. analyses",
        main = "Omics type", xlim=c(0,100))
text(y = (1:length(res2)*1.2)-1.2  , x = res2 + 6, label = res2, pos = 3, cex = 1, col = "black")
dev.off()

```

## Organism

```{r,org1}

org <- x$Organism
org_split <- strsplit(org,", ")
org <- unlist(org_split)
res <- table(org)
res <- res[order(res)]
length(res)
res

par(mar=c(1,1,1,1))
names(res) <- gsub("Homo sapiens","human",names(res))
wordcloud(words = names(res), freq = res, min.freq = 1, 
          max.words=200, random.order=FALSE, rot.per=0.35, 
          colors=brewer.pal(8, "Dark2"), scale=c(4,.5))

par(mar=c(5,12,3,5))
names(res) <- gsub("human","Homo sapiens",names(res))
barplot(tail(res,20),horiz=TRUE,las=1,cex.names = 0.7, xlab="no. analyses",
        main = "Organism", xlim=c(0,200))
grid()

other <- sum(res[1:(nrow(res)-10)])
res2 <- c(other,tail(res,9))
names(res2)[1] <- "Other"
par(mar=c(5,12,3,5))
barplot(res2,horiz=TRUE,las=1,cex.names = 1, xlab="no. analyses",
        main = "Organism", xlim=c(0,200))
text(y = (1:length(res2)*1.2)-1.2  , x = res2+7, label = res2, pos = 3, cex = 1, col = "black")

png("images/organisms1.png",width=400,height=300)
par(mar=c(5,12,3,3))
barplot(res2,horiz=TRUE,las=1,cex.names = 1, xlab="no. analyses",
        main = "Organism", xlim=c(0,200))
text(y = (1:length(res2)*1.2)-1.2  , x = res2 + 15, label = res2, pos = 3, cex = 1, col = "black")
dev.off()

```

## Gene set library

```{r,gsl1}

GSL <-x$Gene.set.library
GSL_split <- strsplit(GSL,", ")
GSL <- unlist(GSL_split)
res <- table(GSL)
res <- res[order(res)]
length(res)
which(names(res)=="Not stated")/sum(res)*100
res

par(mar=c(1,1,1,1))
names(res) <- gsub("Homo sapiens","human",names(res))
wordcloud(words = names(res), freq = res, min.freq = 1, 
          max.words=200, random.order=FALSE, rot.per=0.35, 
          colors=brewer.pal(8, "Dark2"))

par(mar=c(5,12,3,5))
barplot(tail(res,20),horiz=TRUE,las=1,cex.names = 0.7, xlab="no. analyses",
        main = "Gene set library", xlim=c(0,140))
grid()

other <- sum(res[1:(nrow(res)-10)])
res2 <- c(other,tail(res,9))
names(res2)[1] <- "Other"
par(mar=c(5,12,3,5))
barplot(res2,horiz=TRUE,las=1,cex.names = 1, xlab="no. analyses",
        main = "Gene set library", xlim=c(0,140))
text(y = (1:length(res2)*1.2)-1.2  , x = res2+7, label = res2, pos = 3, cex = 1, col = "black")

png("images/genesetlib1.png",width=400,height=300)
par(mar=c(5,12,3,3))
barplot(res2,horiz=TRUE,las=1,cex.names = 1, xlab="no. analyses",
        main = "Gene set library", xlim=c(0,150))
text(y = (1:length(res2)*1.2)-1.2  , x = res2 + 15, label = res2, pos = 3, cex = 1, col = "black")
dev.off()

```

## Gene set version

```{r,gsv1}

GSV <-x$GS.version
res <- table(GSV)
res
res[1]/sum(res)*100

par(mar=c(14,12,3,10))
barplot(tail(res,20),horiz=TRUE,las=1,cex.names = 1, xlab="no. analyses",
        main = "Gene set version defined", xlim=c(0,250))
text(y = (1:length(res)*1.2) - 0.75 , x = res+15, label = res, pos = 3, cex = 1, col = "black")

png("images/genesetvers1.png",width=300,height=150)
par(mar=c(5,5,3,3))
barplot(res,horiz=TRUE,las=1,cex.names = 1, xlab="no. analyses",
        main = "Gene set version defined", xlim=c(0,250))
text(y = (1:length(res)*1.2)-1.2  , x = res + 20, label = res, pos = 3, cex = 1, col = "black")
dev.off()

```

## Statistical test used

```{r,test1}

test <-x$Statistical.test.used
test <- strsplit(test,", ")
test <- unlist(test)
res <- table(test)
res <- res[order(res)]
res[which(names(res)=="Not stated")] / sum(res) * 100
length(res)
res

par(mar=c(1,1,1,1))
wordcloud(words = names(res), freq = res, min.freq = 1, 
          max.words=200, random.order=FALSE, rot.per=0.35, 
          colors=brewer.pal(8, "Dark2"))

par(mar=c(5,12,3,5))
barplot(tail(res,20),horiz=TRUE,las=1,cex.names = 0.7, xlab="no. analyses",
        main = "Test used", xlim=c(0,150))
grid()

other <- sum(res[1:(nrow(res)-10)])
res2 <- c(other,tail(res,9))
names(res2)[1] <- "Other"
par(mar=c(5,12,3,5))
barplot(res2,horiz=TRUE,las=1,cex.names = 1, xlab="no. analyses",
        main = "Test used", xlim=c(0,150))
text(y = (1:length(res2)*1.2)-1.1  , x = res2+10, label = res2, pos = 3, cex = 1, col = "black")

png("images/stattest1.png",width=400,height=300)
par(mar=c(5,12,3,3))
barplot(res2,horiz=TRUE,las=1,cex.names = 1, xlab="no. analyses",
        main = "Test used", xlim=c(0,160))
text(y = (1:length(res2)*1.2)-1.2  , x = res2 + 15, label = res2, pos = 3, cex = 1, col = "black")
dev.off()

```

## FDR Correction

```{r,fdr1}

fdr <-x$FDR.Correction
fdr <- strsplit(fdr,", ")
fdr <- unlist(fdr)
res <- table(fdr)
res <- res[order(res)]
sum(res[which(names(res)!="Yes")])/sum(res)*100
res

par(mar=c(1,1,1,5))
wordcloud(words = names(res), freq = res, min.freq = 1, 
          max.words=200, random.order=FALSE, rot.per=0.35, 
          colors=brewer.pal(8, "Dark2"))

par(mar=c(10,12,3,5))
barplot(tail(res,20),horiz=TRUE,las=1,cex.names = 1, xlab="no. analyses",
        main = "FDR correction performed", xlim=c(0,150))
text(y = (1:length(res)*1.2) - 0.8 , x = res+10, label = res, pos = 3, cex = 1, col = "black")

png("images/fdr1.png",width=300,height=200)
par(mar=c(5,8,3,3))
barplot(res,horiz=TRUE,las=1,cex.names = 1, xlab="no. analyses",
        main = "FDR correction performed", xlim=c(0,160))
text(y = (1:length(res)*1.2)-1.2  , x = res + 15, label = res, pos = 3, cex = 1, col = "black")
dev.off()
```

## App used

```{r,app1}

App <-x$App.used
App_split <- strsplit(App,", ")
App <- unlist(App_split)
res <- table(App)
res <- res[order(res)]
res[which(names(res)=="Not stated")]/sum(res)*100
length(res)
res

par(mar=c(1,1,1,1))
wordcloud(words = names(res), freq = res, min.freq = 1, 
          max.words=200, random.order=FALSE, rot.per=0.35, 
          colors=brewer.pal(8, "Dark2"))

par(mar=c(5,12,3,5))
barplot(tail(res,20),horiz=TRUE,las=1,cex.names = 0.7, xlab="no. analyses",
        main = "App used", xlim=c(0,60))
grid()

other <- sum(res[1:(nrow(res)-10)])
res2 <- c(other,tail(res,9))
names(res2)[1] <- "Other"
par(mar=c(5,12,3,5))
barplot(res2,horiz=TRUE,las=1,cex.names = 1, xlab="no. analyses",
        main = "App used", xlim=c(0,80))
text(y = (1:length(res2)*1.2)-1.1  , x = res2+5, label = res2, pos = 3, cex = 1, col = "black")

png("images/app1.png",width=400,height=300)
par(mar=c(5,12,3,3))
barplot(res2,horiz=TRUE,las=1,cex.names = 1, xlab="no. analyses",
        main = "App used", xlim=c(0,80))
text(y = (1:length(res2)*1.2)-1.2  , x = res2 + 5, label = res2, pos = 3, cex = 1, col = "black")
dev.off()

```

## App version

```{r,appv1}

APV <-x$App.Version
res <- table(APV)
res
res[1]/sum(res)*100

par(mar=c(14,12,3,5))
barplot(tail(res,20),horiz=TRUE,las=1,cex.names = 1, xlab="no. analyses",
        main = "App version defined", xlim=c(0,200))
text(y = (1:length(res)*1.2) - 0.75 , x = res+15, label = res, pos = 3, cex = 1, col = "black")

png("images/appvers1.png",width=300,height=150)
par(mar=c(5,5,3,3))
barplot(res,horiz=TRUE,las=1,cex.names = 1, xlab="no. analyses",
        main = "App version defined", xlim=c(0,200))
text(y = (1:length(res)*1.2)-1.2  , x = res + 15, label = res, pos = 3, cex = 1, col = "black")
dev.off()

```

## Code available

```{r,code1}

code <-x$Code.availability
res <- table(code)
res
res[1]/sum(res)*100

par(mar=c(14,12,3,5))
barplot(tail(res,20),horiz=TRUE,las=1,cex.names = 1, xlab="no. analyses",
        main = "Code availability", xlim=c(0,50))
text(y = (1:length(res)*1.2) - 0.75 , x = res+2, label = res, pos = 3, cex = 1, col = "black")

png("images/code1.png",width=300,height=150)
par(mar=c(5,5,3,3))
barplot(res,horiz=TRUE,las=1,cex.names = 1, xlab="no. analyses",
        main = "Code availability", xlim=c(0,50))
text(y = (1:length(res)*1.2)-1.2  , x = res + 3, label = res, pos = 3, cex = 1, col = "black")
dev.off()

```

## Background gene set

```{r,background1}

BG <-x$Background.gene.set
BG <- gsub("Yes","Yes, correct",BG)
res <- table(BG)
res
sum(res[which(names(res)!="Yes")])/sum(res)*100

par(mar=c(10,12,3,5))
barplot(tail(res,20),horiz=TRUE,las=1,cex.names = 1, xlab="no. analyses",
        main = "Background list specification", xlim=c(0,200))
text(y = (1:length(res)*1.2) - 0.85 , x = res+10, label = res, pos = 3, cex = 1, col = "black")

png("images/bg1.png",width=300,height=200)
par(mar=c(5,9,3,2))
barplot(res,horiz=TRUE,las=1,cex.names = 1, xlab="no. analyses",
        main = "Background list specified", xlim=c(0,220))
text(y = (1:length(res)*1.2)-1.1  , x = res + 18, label = res, pos = 3, cex = 1, col = "black")
dev.off()

```

## Gene lists provided

```{r,genelists1}

GL <-x$Gene.lists.provided
res <- table(GL)
res
sum(res[which(names(res)!="Yes")])/sum(res)*100

par(mar=c(14,12,3,5))
barplot(tail(res,20),horiz=TRUE,las=1,cex.names = 1, xlab="no. analyses",
        main = "Gene lists provided", xlim=c(0,160))
text(y = (1:length(res)*1.2) - 0.75 , x = res+8, label = res, pos = 3, cex = 1, col = "black")

png("images/genelists1.png",width=300,height=150)
par(mar=c(5,5,3,3))
barplot(res,horiz=TRUE,las=1,cex.names = 1, xlab="no. analyses",
        main = "Gene lists provided", xlim=c(0,160))
text(y = (1:length(res)*1.2)-1.2  , x = res + 9, label = res, pos = 3, cex = 1, col = "black")
dev.off()

```

## Methodological flaws

```{r,assumptions1}

ok <- nrow(subset(x,Assumptions.violated=="No"))
ok
bad <- nrow(subset(x,Assumptions.violated!="No"))
bad
ok/sum(bad,ok)*100

ass <-x$Assumptions.violated
ass <- gsub("^No$","None",ass)
ass <- strsplit(ass,", ")
ass <- unlist(ass)
res <- table(ass)
res <- res[order(res)]
res

par(mar=c(1,1,1,1))
wordcloud(words = names(res), freq = res, min.freq = 1, 
          max.words=200, random.order=FALSE, rot.per=0.35, 
          colors=brewer.pal(8, "Dark2"))

par(mar=c(8,12,3,5))
barplot(tail(res,20),horiz=TRUE,las=1,cex.names = 1, xlab="no. analyses",
        main = "Methodological flaws", xlim=c(0,200))
text(y = (1:length(res)*1.2) - 0.9 , x = res+8, label = res, pos = 3, cex = 1, col = "black")

png("images/assumptions1.png",width=400,height=250)
par(mar=c(5,12,3,2))
barplot(res,horiz=TRUE,las=1,cex.names = 1, xlab="no. analyses",
        main = "Methodological flaws", xlim=c(0,220))
text(y = (1:length(res)*1.2)-1.1  , x = res + 18, label = res, pos = 3, cex = 1, col = "black")
dev.off()

```

## Separated or combined

```{r,combined}

comb <-x$Separated.up.and.down
res <- table(comb)
res
sum(res[which(names(res)!="Separated")])/sum(res)*100
res <- res[which(names(res)!="-")]

par(mar=c(10,12,3,5))
barplot(tail(res,20),horiz=TRUE,las=1,cex.names = 1, xlab="no. analyses",
        main = "Separated or combined ORA (differential expression)", xlim=c(0,90))
grid()
text(y = (1:length(res)*1.2) - 0.9 , x = res+5, label = res, pos = 3, cex = 1, col = "black")

png("images/comb1.png",width=300,height=200)
par(mar=c(5,8,3,2))
barplot(res,horiz=TRUE,las=1,cex.names = 1, xlab="no. analyses",
        main = "Separated or combined ORA", xlim=c(0,100))
text(y = (1:length(res)*1.2)-1.1  , x = res + 5, label = res, pos = 3, cex = 1, col = "black")
dev.off()

```

## GSEA parameters

```{r,gseaparam}

param <-x$Parameters
param <- param[which(param!="-")]
param <- gsub(" ","",param)
str(param)

rankspec <- sapply(strsplit(param,","),"[[",1)
rankspec <- gsub("rank=","",rankspec)
rankspec <- table(rankspec)
rankspec

algospec <- sapply(strsplit(param,","),"[[",2)
algospec <- gsub("algo=","",algospec)
algospec <- table(algospec)
algospec

weightspec <- sapply(strsplit(param,","),"[[",3)
weightspec <- gsub("weight=","",weightspec)
weightspec <- table(weightspec)
weightspec

par(mar=c(10,12,3,4))
params <- rbind(rankspec,algospec,weightspec)
rownames(params) <- c("rank method", "test type", "weight type")
colnames(params) <- c("not stated", "stated")
barplot(t(params),horiz=TRUE,las=1,xlim=c(0,30),legend = colnames(params),
        main="GSEA parameter reporting", xlab="no. analyses",
        args.legend = list(x = "topleft", inset = c(0.05, 0.1)))
text(y = (1:nrow(params)*1.2) - 0.7 , x = params[,1]-3, label = params[,1], pos = 3, cex = 1, col = "white")
text(y = (1:nrow(params)*1.2) - 0.7 , x = params[,1]+2, label = params[,2], pos = 3, cex = 1, col = "black")

png("images/gseaparam1.png",width=350,height=250)
par(mar=c(5,8,3,2))
params <- rbind(rankspec,algospec,weightspec)
rownames(params) <- c("rank method", "test type", "weight type")
colnames(params) <- c("not stated", "stated")
barplot(t(params),horiz=TRUE,las=1,xlim=c(0,30),legend = colnames(params),
        main="GSEA parameter reporting", xlab="no. analyses",
        args.legend = list(x = "topleft", inset = c(0.05, 0.1)))
text(y = (1:nrow(params)*1.2) - 0.7 , x = params[,1]-3, label = params[,1], pos = 3, cex = 1, col = "white")
text(y = (1:nrow(params)*1.2) - 0.7 , x = params[,1]+2, label = params[,2], pos = 3, cex = 1, col = "black")
dev.off()

```

## Session information

```{r,session}

sessionInfo()

```