### 2021-11-25 examples.Rmd
more work on examples

---
title: "Example gene set analysis: bringing it together"
author: "Mark Ziemann & Kaumadi Wijesooriya"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
    fig_width: 7
    fig_height: 5
theme: cosmo
---

Source: https://github.com/markziemann/SurveyEnrichmentMethods

## Intro

Previously we analysed five datasets below and looked at the differences between FCS and ORA methods for gene set enrichment.

| Dataset | SRA accesion | genes in annotation set | genes detected | genes differentially expressed |
| :---: | :---: | :---: | :---: | :---: | 
| 1. | SRP128998 | 39297 | 15635 | 3472 |
| 2. | SRP038101 | 39297 | 13926 | 3589 | 
| 3. | SRP096178 | 39297 | 15477 | 9488 |
| 4. | SRP038101 | 39297 | 15607 | 5150 | 
| 5. | SRP247621 | 39297 | 14288 | 230 |
| 6. | SRP253951 | 39297 | 15182 | 8588 |
| 7. | SRP068733 | 39297 | 14255 | 7365 |

```{r,begin}

suppressPackageStartupMessages({
library("kableExtra")
library("beeswarm")
})

```

## Summarise

Here I'm summarising the results of the five example data sets.

```{r,getdata,fig.width=7,fig.height=5}

exdat=NULL
ex1dat <- readRDS("ex1dat.rds")
ex2dat <- readRDS("ex2dat.rds")
ex3dat <- readRDS("ex3dat.rds")
ex4dat <- readRDS("ex4dat.rds")
ex5dat <- readRDS("ex5dat.rds")
ex6dat <- readRDS("ex6dat.rds")
ex7dat <- readRDS("ex7dat.rds")

exdat <- rbind(ex1dat,ex2dat,ex3dat,ex4dat,ex5dat,ex6dat,ex7dat)
rownames(exdat) <- c("SRP128998", "SRP038101", "SRP096178","SRP038101","SRP247621","SRP253951","SRP068733")

exdat %>% kbl(caption="jaccard index values across seven RNA-seq studies") %>% kable_paper("hover", full_width = F)

```

```{r,summary_stats}

means <- colMeans(exdat)
medians <- colMedians(exdat)
sds <- sd(exdat)
summstats <- rbind(means,medians,sds)
summstats %>% kbl(caption="summary statistics of different enrichment analysis comparisons") %>% kable_paper("hover", full_width = F)

```

Here let's plot the jaccard scores.
First as a boxplot and then overlay the dataset studies over the top.

```{r,jacbox}

cols <- c("#CC79A7","#56B4E9","#E69F00","#F0E442","#009E73","#0072B2","#D55E00")
coll <- list(cols, cols, cols, cols, cols)
par(mar=c(5,10,3,1))
boxplot(list(exdat[,5],exdat[,4],exdat[,3],exdat[,2],exdat[,1]),names = rev(colnames(exdat)),horizontal = TRUE, las=1 , col="white", cex=1.5,pch=19, xlab = "Jaccard index")
beeswarm(list(exdat[,5],exdat[,4],exdat[,3],exdat[,2],exdat[,1]),labels = rev(colnames(exdat)),horizontal = TRUE, las=1 , pwcol = coll,cex=1.5,pch=19, xlab = "Jaccard index",add = TRUE)
legend("bottomright", legend=c("SRP128998", "SRP038101", "SRP096178","SRP038101","SRP247621","SRP253951","SRP068733"), pch=19,  col=cols, cex=1,pt.cex = 1.5)

pdf("images/jacbox1.pdf",width=5,height=4)
par(mar=c(5,9,3,2))
boxplot(list(exdat[,5],exdat[,4],exdat[,3],exdat[,2],exdat[,1]),names = rev(colnames(exdat)),horizontal = TRUE, las=1 , col="white", cex=1.5,pch=19, xlab = "Jaccard index")
beeswarm(list(exdat[,5],exdat[,4],exdat[,3],exdat[,2],exdat[,1]),labels = rev(colnames(exdat)),horizontal = TRUE, las=1 , pwcol = coll,cex=1.5,pch=19, xlab = "Jaccard index",add = TRUE)
legend("topleft", legend=c("SRP128998", "SRP038101", "SRP096178","SRP038101","SRP247621","SRP253951","SRP068733"), pch=19,  col=cols, cex=1,pt.cex = 1.5)
dev.off()

png("images/jacbox1.png")
par(mar=c(5,9,3,1))
boxplot(list(exdat[,5],exdat[,4],exdat[,3],exdat[,2],exdat[,1]),names = rev(colnames(exdat)),horizontal = TRUE, las=1 , col="white", cex=1.5,pch=19, xlab = "Jaccard index")
beeswarm(list(exdat[,5],exdat[,4],exdat[,3],exdat[,2],exdat[,1]),labels = rev(colnames(exdat)),horizontal = TRUE, las=1 , pwcol = coll,cex=1.5,pch=19, xlab = "Jaccard index",add = TRUE)
legend("bottomright", legend=c("SRP128998", "SRP038101", "SRP096178","SRP038101","SRP247621","SRP253951","SRP068733"), pch=19,  col=cols, cex=1.2,pt.cex = 1.5)
dev.off()

```

Now plot as a simple bargraph

```{r,jacbar}

mymean <- colMeans(exdat)
mysd <- apply(exdat,2,sd)

par(mar=c(5,10,3,1))
barplot(rev(colMeans(exdat)),xlab="mean Jaccard index",horiz = TRUE, las =1, xlim=c(0,.7) )

pdf("images/jacbar1.pdf",width=4,height=4)
par(mar=c(5,10,3,1))
barplot(rev(colMeans(exdat)),xlab="mean Jaccard index",horiz = TRUE, las =1, xlim=c(0,.7) )
dev.off()

png("images/jacbar1.png")
par(mar=c(5,10,3,1))
barplot(rev(colMeans(exdat)),xlab="mean Jaccard index",horiz = TRUE, las =1, xlim=c(0,.7) )
dev.off()

```

Now plot as a beeswarm.

```{r,jacswarm}

exdat <- rbind(exdat,"mean"=colMeans(exdat))

cols <- c("#CC79A7","#56B4E9","#E69F00","#F0E442","#009E73","black")
coll <- list(cols, cols, cols, cols, cols)
par(mar=c(5,10,3,1))
beeswarm(list(exdat[,5],exdat[,4],exdat[,3],exdat[,2],exdat[,1]),labels = rev(colnames(exdat)),horizontal = TRUE, las=1 , pwcol = coll,cex=2,pch=19, xlab = "Jaccard index")
legend("bottomright", legend=c("SRP128998", "SRP038101", "SRP096178","SRP038101","SRP247621","mean"), pch=19,  col=cols, cex=1,pt.cex = 1.5)

pdf("images/jacswarm1.pdf",width=6,height=5)
par(mar=c(5,9,3,1))
beeswarm(list(exdat[,5],exdat[,4],exdat[,3],exdat[,2],exdat[,1]),labels = rev(colnames(exdat)),horizontal = TRUE, las=1 , pwcol = coll,cex=2,pch=19, xlab = "Jaccard index")
legend("bottomright", legend=c("SRP128998", "SRP038101", "SRP096178","SRP038101","SRP247621","mean"), pch=19,  col=cols, cex=1.0, pt.cex = 1.7)
dev.off()

png("images/jacswarm1.png")
par(mar=c(5,10,3,1))
beeswarm(list(exdat[,5],exdat[,4],exdat[,3],exdat[,2],exdat[,1]),labels = rev(colnames(exdat)),horizontal = TRUE, las=1 , pwcol = coll,cex=2,pch=19, xlab = "Jaccard index")
legend("bottomright", legend=c("SRP128998", "SRP038101", "SRP096178","SRP038101","SRP247621","mean"), pch=19,  col=cols, cex=1,pt.cex = 1.5)
dev.off()

rbind(exdat,mysd) %>% kbl() %>% kable_paper("hover", full_width = F)

```

## Session information

```{r,session}

sessionInfo()

```