### 2021-12-01 replication/PMC6463127.Rmd
PMC6463127.Rmd

---
title: "PMC6463127: A replication study"
author: "Mark Ziemann & Kaumadi Wijesooriya"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
    fig_width: 7
    fig_height: 5
theme: cosmo
---

Source: https://github.com/markziemann/SurveyEnrichmentMethods

## Intro

```{r,begin}

suppressPackageStartupMessages({
library("getDEE2")
library("DESeq2")
library("clusterProfiler")
library("org.Hs.eg.db")
library("mitch")
library("kableExtra")
library("eulerr")
})

```

In [Kaumadi et al 2021](https://www.biorxiv.org/content/10.1101/2021.09.06.459114v1) we
showed that major problems plague functional enrichment analysis in many journal articles.

Here I will do my best to reproduce and replicate the findings of some studies.
These were selected on the following criteria:

* Included in the 2019 group that was studies in Kaumadi et al 2021.

* Human study

* RNA-seq gene expression

* Performed DAVID gene set analysis

* Provided gene list

From this subset, four studies were selected

| Study | Analytical issues |
| --- | --- |
| PMC6349697 | Background |
| PMC6425008 | Background, FDR |
| PMC6463127 | Background, FDR |
| PMC6535219 | Background |

In this document I will be seeing if I can reproduce the same findings by using the same
tool with the gene list provided in the article.
Also I will correct the method wherever possible.
This means using the correct background gene list, using FDR control and performing 
separate analysis of up and down-regulated gene lists.

The study examined in this document is PMC6463127: 
[Chai et al, 2019](https://pubmed.ncbi.nlm.nih.gov/30832348/).

The focus of this study is to compare the RNA expression profiles of parathyroid adenomas with normal parathyroid gland tissue.
They found 247 differentially expressed genes; 45 up-regulated, 202 down-regulated.

They used DAVID for GO/KEGG analysis.
They did not perform FDR correction of enrichment results.

Here are the upregulated and downregulated GO terms and the downregulated KEGG pathways. 
(Presuming there were no significantly upregulated KEGG pathways.)

```{r,orig1}

orig1 <- read.table("PMC6463127_orig_up.tsv",header=TRUE,sep="\t")

orig1 %>% kbl(caption="upregulated GO terms shown in the article") %>% kable_paper("hover", full_width = F)

orig2 <- read.table("PMC6463127_orig_dn.tsv",header=TRUE,sep="\t")

orig2 %>% kbl(caption="down-regulated GO terms shown in the article") %>% kable_paper("hover", full_width = F)

orig3 <- read.table("PMC6463127_orig_keggdn.tsv",header=TRUE,sep="\t")

orig3 %>% kbl(caption="down-regulated KEGG pathways shown in the article") %>% kable_paper("hover", full_width = F)

```

## Try to reproduce

So let's see if I can reproduce their result using DAVID without a background gene list.
The study provided the differentially regulated gene lists in Table S2

Here are the significant GO terms for the up and down directions, along with the downregulated KEGG pathways.

```{r,repl1}

repl1up <- read.table("PMC6463127_repl_up.tsv",header=TRUE,sep="\t")

repl1up %>% kbl(caption="Upregulated GO terms obtained using article gene list without background") %>% kable_paper("hover", full_width = F)

repl1dn <- read.table("PMC6463127_repl_dn.tsv",header=TRUE,sep="\t")

repl1dn %>% kbl(caption="Downregulated GO terms obtained using article gene list without background") %>% kable_paper("hover", full_width = F)

repl1kegg <- read.table("PMC6463127_repl_kegg.tsv",header=TRUE,sep="\t")

repl1kegg %>% kbl(caption="Downregulated KEGG pathways obtained using article gene list without background") %>% kable_paper("hover", full_width = F)

```

There is a supplement with all expressed genes (Supplementary Table S2) which I will use as a background.
To get this to work, I will convert the gene symbols to Ensembl identifiers.
There were only ~8000 expressed genes which is a bit low.

```{r,bg}

bgg <- readLines("PMC6463127_bgg.txt")
head(bgg)
length(bgg)

geneinfo <- read.table("http://dee2.io/data/hsapiens/hsa_gene_info.tsv",header=TRUE)

bg <- geneinfo[which(geneinfo$GeneSymbol %in% bgg),"GeneID"]
bg <- bg[which(!is.na(bg))]
bg <- unique(bg)
length(bg)
writeLines(bg,"PMC6463127_bge.txt")

```

## Try to reproduce with corrected background gene list

Then I ran DAVID again using the correct background. 
Here is the result of the upregulated genes.

```{r,replbg1up}

repl2up <- read.table("PMC6463127_repl_up2.tsv",header=TRUE,sep="\t")

repl2up %>% kbl(caption="Upregulated GO terms obtained using article gene list with a background") %>% kable_paper("hover", full_width = F)

repl2dn <- read.table("PMC6463127_repl_dn2.tsv",header=TRUE,sep="\t")

repl2dn %>% kbl(caption="Downregulated GO terms obtained using article gene list with a background") %>% kable_paper("hover", full_width = F)

repl2kegg <- read.table("PMC6463127_repl_kegg2.tsv",header=TRUE,sep="\t")

repl2kegg %>% kbl(caption="Downregulated KEGG pathways obtained using article gene list with a background") %>% kable_paper("hover", full_width = F)

```

## What about a replication using an R script

Here I'm using the enrichGO function to analyse the data.
The algorithm is slightly different and the gene sets might be a different version.
First GO biological processes.

```{r,repl2_go}

# from supp table 2
upg <- readLines("PMC6463127_up.txt")
dng <- readLines("PMC6463127_dn.txt")

upe <- geneinfo[which(geneinfo$GeneSymbol %in% upg),"GeneID"]
dne <- geneinfo[which(geneinfo$GeneSymbol %in% dng),"GeneID"]

# up
go_up <- enrichGO(gene = upe,  keyType = "ENSEMBL", universe = bg,
    OrgDb = org.Hs.eg.db, ont = "ALL", pAdjustMethod = "BH", pvalueCutoff  = 0.05,
    qvalueCutoff  = 1, readable = TRUE)

go_up <- data.frame(go_up)
nrow(go_up)

go_up <- subset(go_up,qvalue<0.05)
nrow(go_up)

go_up[,c(1,3:7)] %>% kbl(caption="Upregulated GO terms obtained using clusterProfiler with the correct background") %>% kable_paper("hover", full_width = F)

# dn
go_dn <- enrichGO(gene = dne,  keyType = "ENSEMBL", universe = bg,
    OrgDb = org.Hs.eg.db, ont = "ALL", pAdjustMethod = "BH", pvalueCutoff  = 0.05,
    qvalueCutoff  = 1, readable = TRUE)

go_dn <- data.frame(go_dn)
nrow(go_dn)

go_dn <- subset(go_dn,qvalue<0.05)
nrow(go_dn)

go_dn[,c(1,3:7)] %>% kbl(caption="Downregulated GO terms obtained using clusterProfiler with the correct background") %>% kable_paper("hover", full_width = F)

```

Now KEGG analysis but need to convert to entrez first

```{r,repl2_kegg}

dn_entrez <- unlist(mget(dne, org.Hs.egENSEMBL2EG, ifnotfound = NA))

bg_entrez <- unlist(mget(bg, org.Hs.egENSEMBL2EG, ifnotfound = NA))

kegg <- enrichKEGG(gene = dn_entrez, universe = bg_entrez,
    organism = "hsa", pAdjustMethod = "BH", pvalueCutoff  = 0.05,
    qvalueCutoff  = 1)

kegg <- as.data.frame(kegg)

kegg[,c(2:7)] %>% kbl(caption="Clusterprofiler KEGG results obtained using the correct background") %>% kable_paper("hover", full_width = F)

```

## Are the conclusions of the study supported?

In the article, they obtained 45 upregulated and 78 downregulated GO terms, along with 5 downregulated KEGG pathways.

The authors don't specifically draw any conclusions from the enrichment analysis, but let the reader come to their
own conclusions on the basis of the results tables presented.

Anyway, we can look at whether the top results were confirmed with a corrected method.

I repeated the analysis without a background and received the same upregulated GO terms as the article 
(16 terms with p<0.05), but of these, only 5 gave an FDR<0.05.

When a correct background gene list was used, only 2 GOs were upregulated with FDR<0.05.
Basically, there were no significantly upregulated GO BP or MF terms when using a correct background and FDR control. 
Using a clusterProfiler script, I got 9 upregulated GOs, 8 of which were different to the DAVID results.

It was the same story with the downregulated GOs, but after using a background gene list
the p-values were larger and the ranks of the terms were changed.

I was able to repeat the KEGG results without a specialised background.
When using the appropriate background, the KEGG results were similar.

In summary the results shown in the article were described as significant, but as
authors did not perform FDR correction, they are not reliable.

Incidentally, the hypergeometric test used in clusterProfiler found many more ontologies
and KEGG pathways that were statistically significant.

## Session information

```{r,session}

sessionInfo()

```