### 2021-12-03 citation_fetching.Rmd
using JSON to retrieve results

---
title: "Fetch PMC Citation Info"
author: "Mark Ziemann"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    fig_width: 7
    fig_height: 7
theme: cosmo
---

## Intro

Here we are going to fetch the number of citations as well as the publication date for every PMCID.

The code below fetches the number of times that A PMC has been cited.

The output format is

1. PMCID

2. PMID

3. Number of cites

4. Pubdate

```{r,library}

library("reutils")
library("XML")
library("kableExtra")
library("anytime")

```

Here are the functions.

You can see that there are tryCatch expressions used to manage the occasional error that are raised by 
eutils queries which I can't reproduce nor debug. 
So for those queries I just attempt again.

```{r,functions}

cites <- function(PMID) {
  tryCatch(
    expr = {
        CITES <- elink(uid=PMID , dbFrom="pubmed", linkname = "pubmed_pubmed_citedin" )
    },
    error = function(e){ 
        CITES <- elink(uid=PMID , dbFrom="pubmed", linkname = "pubmed_pubmed_citedin" )
    },
    finally = {
      return(CITES)
    }
)}

get_cites <- function(PMID) {
  CITES <- cites(PMID)
  N_CITES <- length(CITES[1])
  return(c(PMID,N_CITES))
}

# test the function
PMID=30763910
res <- get_cites(PMID = PMID)
res

```

## Get cites for small group

Now to get results for a few PMC IDs to test the functions are working OK.

```{r,test1}

PMCIDS <- c("PMC6317205", "PMC6317219", "PMC6318854", "PMC6318897", "PMC6322516", "PMC6323737", "PMC6323741", "PMC6325784", "PMC6325795", "PMC6325889", "PMC6326702", "PMC6327033",
            "PMC6327913", "PMC6328174", "PMC6328465", "PMC6328499", "PMC6329100", "PMC6329100", "PMC6329125", "PMC6329125", "PMC6329193", "PMC6329348")

PMCIDS <- gsub("PMC","",PMCIDS)
es <- esummary(PMCIDS, "pmc",retmode = "json")

es <- content(es, "parsed")
es <- es[2]
es <- es[[1]]
es <- es[2:length(es)]

pmid <- unlist(lapply(es,function(x){
  x$articleids[1,2]
}))

pubdate <- unlist(lapply(es,function(x){
  x$pubdate
}))

df <- data.frame(PMCIDS,pmid,pubdate)

ncites <- t(sapply(df[,2],get_cites))
ncites

df$ncites <- ncites[,2]
head(df)
df <- df[,c(1,2,4,3)]
colnames(df) <- c("PMCID","PMID","NumCites","Pubdate")
df$PMCID <- gsub("^","PMC",df$PMCID)
df$Pubdate <- anydate(df$Pubdate)
df$NumCites <- as.numeric(df$NumCites)
str(df)

df %>% kbl %>% kable_paper("hover", full_width = F)

plot(df$Pubdate,df$NumCites)
mylm <- lm(df$NumCites ~ df$Pubdate)

mylm

```

Looks like it works :)

Except reutils fails for articles with any special characters in there.

## Get cites for all

Now to get a bunch of these.

```{r,getcites}

x <- read.table("data/PMC_2019-analysis.tsv",sep="\t",header=TRUE)
head(x)
PMCIDS <- unique(x$Pubmed.Central.ID)

mycitesfunc <- function(PMCIDS) {
  PMCIDS <- gsub("PMC","",PMCIDS)
  es <- esummary(PMCIDS, "pmc",retmode = "json")
  es <- content(es, "parsed")
  es <- es[2]
  es <- es[[1]]
  es <- es[2:length(es)]
  pmid <- unlist(lapply(es,function(x){
    x$articleids[1,2]
  }))
  pubdate <- unlist(lapply(es,function(x){
    x$pubdate
  }))
  df <- data.frame(PMCIDS,pmid,pubdate)
  ncites <- t(sapply(df[,2],get_cites))
  df$ncites <- ncites[,2]
  df <- df[,c(1,2,4,3)]
  colnames(df) <- c("PMCID","PMID","NumCites","Pubdate")
  df$PMCID <- gsub("^","PMC",df$PMCID)
  df$Pubdate <- anydate(df$Pubdate)
  df$NumCites <- as.numeric(df$NumCites)
  return(df)
}

df1 <- mycitesfunc(PMCIDS[1:500])
df2 <- mycitesfunc(PMCIDS[501:1000])
df3 <- mycitesfunc(PMCIDS[1001:1500])

df <- rbind(df1,df2,df3)

df %>% kbl %>% kable_paper("hover", full_width = F)

plot(df$Pubdate,df$NumCites)
mylm <- lm(df$NumCites ~ df$Pubdate)

mylm

saveRDS(df,"cites_data.rds")

```

## Session information

```{r,sessioninfo}

sessionInfo()

```