### 2021-12-07 reproduce_analysis_and_manuscript.Rmd
finishing touches

---
title: "Reproduce the analysis and regenerate the manuscript"
author: "Mark Ziemann & Kaumadi Wijesooriya"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    fig_width: 7
    fig_height: 5
theme: cosmo
---

Source: https://github.com/markziemann/SurveyEnrichmentMethods

## Intro

This script is designed to enable the complete analysis described in the article 
to be replicated with a single command.

By knitting this Rmd, it will reanalyse the data and regenerate all figures of
the manuscript.

First of all, it is a good idea to check whether all of the requisite 
dependancies are installed.

Ubuntu users may need to install `librsvg2-dev` and `libv8-dev` to get `DiagrammeR` to work.

```{r,libraries}

# for figure 2 and 3
library("wordcloud")
library("vioplot")
library("reutils")
library("XML")
library("kableExtra")
library("anytime")

# for figure 4
library("getDEE2") 
library("DESeq2")
library("clusterProfiler")
library("mitch")
library("eulerr")
library("beeswarm")

# for figure 1 / manuscript draft
library("rmdformats")
library("DiagrammeR")
library("DiagrammeRsvg")
library("rsvg")

```

## Execute the scripts

Here will re-run all of the analysis and re-build the manuscript draft.

```{r,run_all}

# Figure 2
rmarkdown::render("analysis_code/analysis1.Rmd")

# Figure 3
rmarkdown::render("analysis_code/analysis2.Rmd")

# uncomment the below line in order to update the number of accumulated
# citations for each article considered
#rmarkdown::render("analysis_code/citation_fetching.Rmd")

# Figure 4
rmarkdown::render("examples/examples.Rmd")

# Figure 1 and assemble a manuscript draft
rmarkdown::render("docs/manuscript.Rmd")

```