### 2022-01-12 analysis_code/analysis2.Rmd
fixed minor issue with journal name mismatches betwen scimago and nlm

---
title: "PMC Survey Analysis Part 2"
author: "Mark Ziemann & Kaumadi Wijesooriya"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    fig_width: 7
    fig_height: 5
theme: cosmo
---

Source: https://github.com/markziemann/SurveyEnrichmentMethods

## Intro

Here we are performing the analysis that is described in Figure 3 of the manuscript.
This involves calculating an analysis score for a 1500 articles from 2019.
We then identify the best/worst journals, and correlate analysis scores with journal and article level metrics.

```{r,begin}

knitr::opts_chunk$set(fig.width=7, fig.height=5)

library("wordcloud")
library("vioplot")
library("ggplot2")
library("ggExtra")
library("europepmc")

```

## Overview of included and excluded analyses

```{r,exclude1}

x <- read.table("../data/PMC_2019-analysis.tsv",header=TRUE,fill=TRUE,sep="\t")
y <- read.table("../data/QC-analysis2.tsv",header=TRUE,fill=TRUE,sep="\t")
y <- y[,1:16]
x <- x[which(! x$Pubmed.Central.ID %in% y$Pubmed.Central.ID),]
x <- rbind(x,y)

head(x)
colnames(x)
dim(x)

exclude <- subset(x,x$GS.version=="EXCLUDE")
nrow(exclude)
length(unique(exclude$Pubmed.Central.ID))

x <- subset(x,x$GS.version!="EXCLUDE")
nrow(x)
length(unique(x$Pubmed.Central.ID))

write.table(x,file="TableS2.tsv",quote=FALSE,sep="\t",col.names = TRUE,row.names = FALSE)

```

## Score

Here I am proposing a scoring scheme.
If there is information missing from the article, or basic mistakes are made, then a point is deducted.
If the article goes over and above the basic reproducibility, then they are awarded a point.

Gene set library origin not stated = -1

Gene set library version not stated = -1

Stat test not stated = -1

No stat test conducted = -1

No FDR correction conducted = -1

App used not stated = -1

App version not stated = -1

Background list not defined = -1

Inappropriate background list used = -1

Code availability = +1

Gene lists provided = +1

```{r,score}

score <- function(r){
  r[is.na(r)] <- 0
  SCORE=0
  # gene set lib
  if(r["Gene.set.library"]=="Not stated"){
    SCORE=SCORE-1
  }
  # GS version
  if(r["GS.version"]=="No"){
    SCORE=SCORE-1
  }
  # stat test
  if(r["Statistical.test.used"]=="No test"){
    SCORE=SCORE-1
  }
  if(r["Statistical.test.used"]=="Not stated"){
    SCORE=SCORE-1
  }
  # FDR
  if(r["FDR.Correction"]!="Yes"){
    SCORE=SCORE-1
  }
  # app used
  if(r["App.used"]=="Not stated"){
    SCORE=SCORE-1
  }
  # App version
  if(r["App.Version"]=="No"){
    SCORE=SCORE-1
  }
  # Code availability
  if(r["Code.availability"]=="Yes"){
    SCORE=SCORE+1
  }
  # Background
  if(r["Background.gene.set"]=="Not stated"){
    SCORE=SCORE-1
  }
  if(r["Background.gene.set"]=="No"){
    SCORE=SCORE-1
  }
  if(r["Background.gene.set"]=="Stated, but incorrect"){
    SCORE=SCORE-1
  }
  # gene list
  if(r["Gene.lists.provided"]=="Yes"){
    SCORE=SCORE+1
  }
  return(SCORE)
}

scores <- apply(X = x,MARGIN = 1, FUN = score)
barplot(table(scores), xlab="analysis score",ylab="frequency")
x$scores <- scores
length(which(scores>0))

mean(x$scores)
sd(x$scores)
median(x$scores)
length(scores)
which(scores>0)

table(scores)

dir.create("images")
png("images/hist1.png",width=300,height=200)
par(mar=c(5,5,2,1))
barplot(table(scores), xlab="analysis score",ylab="frequency",ylim=c(0,600))
text(x= 1:length(table(scores))* 1.2 -0.5, y = table(scores), label = table(scores), pos = 3, cex = 1, col = "black")
dev.off()

pdf("images/hist1.pdf",width=5,height=4)
par(mar=c(5,5,2,1))
barplot(table(scores), xlab="analysis score",ylab="frequency",ylim=c(0,600))
text(x= 1:length(table(scores))* 1.2 -0.5, y = table(scores), label = table(scores), pos = 3, cex = 1, col = "black")
dev.off()

scores2 <- round(table(scores)/sum(table(scores))*100,1)
scores2

png("images/hist2.png",width=300,height=200)
par(mar=c(5,5,2,1))
barplot(scores2, xlab="analysis score",ylab="frequency",ylim=c(0,50))
text(x= 1:length(scores2)* 1.2 -0.5, y = scores2, label = scores2, pos = 3, cex = 0.9, col = "black")
dev.off()

pdf("images/hist2.pdf",width=5,height=4)
par(mar=c(5,5,2,1))
barplot(scores2, xlab="analysis score",ylab="frequency",ylim=c(0,40))
text(x= 1:length(scores2)* 1.2 -0.5, y = scores2, label = scores2, pos = 3, cex = 0.9, col = "black")
dev.off()

```

## Journal rank

There is a problem that the journal names from NLM/PUBMED CENTRAL do not exactly match those given by SciMago.
This results in some journals being dropped from the SJR correlation analysis.
To avoid this, I will use the ISSN numbers which should be unique and non-ambiguous identifiers.
This is obtained with a query of Europe PMC with the europepmc package.

```{r,journal rank}

jmetrics <- read.table("../data/scimagojr_2020.csv",sep=";",header=TRUE,dec = ",")
jmetrics$Title <- toupper(jmetrics$Title)
head(jmetrics,3)
dim(jmetrics)

getissn <- function(pmc){
  Sys.sleep(1)
  pmcres <- europepmc::epmc_search(pmc)
  pmcissn <- pmcres$journalIssn
  pmcissn <- gsub("-","",pmcissn)
  return(pmcissn)
}

issn <- sapply(x$Pubmed.Central.ID,getissn)
issn2 <- lapply(issn,function(x) { unlist(strsplit(x,"\\; ")) } )
l <- lapply(issn2,length)
l <- unlist(l)
lnames <- names(l)
pmcnames <- unlist(sapply(1:length(l), function(i) { rep(lnames[i],l[i]) }  ))
issn2 <- unname(unlist(issn2))
issn2 <- gsub("-","",issn2)
pmcissn <- data.frame(pmcnames,issn2)
xissn <- merge(x,pmcissn,by.x="Pubmed.Central.ID",by.y="pmcnames")

xm <- merge(xissn,jmetrics,by.x="issn2",by.y="Issn")

```

## Analysis scores vs Journal rank

Here I will be making a scatter plot of the analysis scores vs the SJR.

I will also attempt to make the same scatterplot with marginal histograms according to (https://stackoverflow.com/questions/8545035/scatterplot-with-marginal-histograms-in-ggplot2)

```{r,scoresvsjrank1}

tail(xm$Journal)

hist(xm$SJR,xlab="SJR",main="SJR distribution")

mylm1 <- lm (xm$scores ~ xm$SJR)
plot(xm$SJR,xm$scores, col = rgb(red=0,blue=0,green=0,alpha=0.2), pch=19, bty="n",
  xlab="SJR",ylab="score",main="Analysis scores vs. journal metrics")
abline(mylm1,col="red")

cor.test(xm$scores,xm$SJR,method="pearson")
cor.test(xm$scores,xm$SJR,method="spearman")

png("images/score_sjr1.png",width=350,height=300)
par(mar=c(5,5,3,1))
plot(xm$SJR,xm$scores, col = rgb(red=0,blue=0,green=0,alpha=0.2), pch=19,bty="n",
  xlab="SJR",ylab="score",main="Analysis scores vs. journal metrics")
abline(mylm1,col="red")
grid()
dev.off()

pdf("images/score_sjr1.pdf",width=4,height=3)
par(mar=c(4,4,3,2))
plot(xm$SJR,xm$scores, col = rgb(red=0,blue=0,green=0,alpha=0.2), pch=19,bty="n",
  xlab="SJR",ylab="score",main="Analysis scores vs. journal metrics")
abline(mylm1,col="red")
grid()
dev.off()

df <- data.frame("SJR" = xm$SJR, "score" = xm$scores )
p <- ggplot(df, aes(SJR, score)) + 
  geom_point(alpha=0.1, size = 3) + 
  stat_smooth(method = lm,fullrange=TRUE) +
  theme_classic()
ggExtra::ggMarginal(p, type = "histogram")

pdf("images/score_sjr2.pdf",width=4,height=3)
p <- ggplot(df, aes(SJR, score)) + 
  geom_point(alpha=0.1, size = 3) + 
  stat_smooth(method = lm,fullrange=TRUE) +
  theme_classic()
ggExtra::ggMarginal(p, type = "histogram")
dev.off()

```

## Best and worst journals

Minimum of 5 analyses.

```{r,goodbadugly,fig.height=5,fig.width=6}

x2 <-x[,c("Journal","scores")]
tab <- table(x2$Journal)
tab <- tab[which(tab>=5)]
x2 <- x2[which(x2$Journal %in% names(tab)),]
dim(x2)
jres <- aggregate(scores ~ Journal, x2 ,mean)
jsd <- aggregate(scores ~ Journal, x2 ,sd)
jres$sd <- jsd$scores
jres <- jres[order(jres$scores),]
dim(jres)
head(jres)

par(mar=c(5,12,3,1))
barplot(tail(jres$scores,20),names.arg = tail(jres$Journal,20),horiz=TRUE,las=1,cex.names = 0.7, xlab="mean score",
        main = "Highest scoring",xlim=c(-4,0))
grid()

png("images/score_highest1.png",width=300,height=350)
par(mar=c(5,12,3,1))
barplot(tail(jres$scores,20),names.arg = tail(jres$Journal,20),horiz=TRUE,las=1,cex.names = 0.7, xlab="mean score",
        main = "Highest scoring",xlim=c(-4,0))
grid()
dev.off()

par(mar=c(5,12,3,1))
barplot(head(jres$scores,20),names.arg = head(jres$Journal,20),horiz=TRUE,las=1,cex.names = 0.7, xlab="mean score",
        main = "Lowest scoring journals",xlim=c(-5,0))
grid()

png("images/score_lowest1.png",width=300,height=350)
par(mar=c(5,12,3,1))
barplot(head(jres$scores,20),names.arg = head(jres$Journal,20),horiz=TRUE,las=1,cex.names = 0.7, xlab="mean score",
        main = "Lowest scoring",xlim=c(-5,0))
grid()
dev.off()

pdf("images/score_highest1.pdf",width=3,height=4.5)
par(mar=c(5,8,3,1))
barplot(tail(jres$scores,20),names.arg = tail(jres$Journal,20),horiz=TRUE,las=1,cex.names = 0.7, xlab="mean score",
        main = "Highest scoring",xlim=c(-4,0))
grid()
dev.off()

pdf("images/score_lowest1.pdf",width=3,height=4.5)
par(mar=c(5,8,3,1))
barplot(head(jres$scores,20),names.arg = head(jres$Journal,20),horiz=TRUE,las=1,cex.names = 0.7, xlab="mean score",
        main = "Lowest scoring",xlim=c(-5,0))
grid()
dev.off()

j <- unique(x2$Journal)
jscores <- sapply(j,function(jj) {
 x2[which(x2$Journal == jj ),2]  
})

jscores <- jscores[order(unlist(lapply(jscores,mean)))]

par(mar=c(5,12,3,1))
vioplot(tail(jscores,20),horizontal = TRUE,las=1,cex.axis=0.75,main="Highest scoring journals")

par(mar=c(5,12,3,1))
vioplot(head(jscores,20),horizontal = TRUE,las=1,cex.axis=0.75,main="Lowest scoring journals")

```

## Mean analysis scores vs Journal rank

```{r,scoresvsjrank2,fig.height=5,fig.width=6}

x2 <-xm[,c("Journal","SJR","scores")]
tab <- table(x2$Journal)
tab <- tab[which(tab>=5)]
x2 <- x2[which(x2$Journal %in% names(tab)),]
dim(x2)
jres <- aggregate(scores ~ Journal, x2 ,mean)
jsjr <- aggregate(SJR ~ Journal, x2 ,mean)
jres$SJR <- jsjr$SJR
jres <- jres[order(jres$scores),]
dim(jres)
head(jres)

mylm2 <- lm (jres$scores ~ jres$SJR)
plot(jres$SJR,jres$scores,col = rgb(red=0,blue=0,green=0,alpha=0.2) , pch=19,bty="n",
  xlab="SJR",ylab="mean score",main="Mean analysis score vs. journal metrics")
abline(mylm2,col="red")

png("images/mean_sjr1.png",width=350,height=300)
par(mar=c(5,5,3,1))
plot(jres$SJR,jres$scores,  col = rgb(red=0,blue=0,green=0,alpha=0.2) , pch=19,bty="n",
  xlab="SJR",ylab="mean score",main="Mean analysis score vs. journal metrics")
abline(mylm2,col="red")
dev.off()

cor.test(jres$scores,jres$SJR,method="pearson")

cor.test(jres$scores,jres$SJR,method="spearman")

pdf("images/mean_sjr1.pdf",width=3.5,height=3)
par(mar=c(5,5,3,1))
plot(jres$SJR,jres$scores,  col = rgb(red=0,blue=0,green=0,alpha=0.2) , pch=19,bty="n",
  xlab="SJR",ylab="mean score",main="Mean analysis score vs. journal metrics")
abline(mylm2,col="red")
dev.off()

df <- data.frame("SJR" = jres$SJR, "mean score" = jres$scores )
p <- ggplot(df, aes(SJR, mean.score)) +
  geom_point(alpha=0.3, size = 3) +
  stat_smooth(method = lm,fullrange=TRUE) +
  theme_classic()
ggExtra::ggMarginal(p, type = "histogram")

pdf("images/mean_sjr2.pdf",width=4,height=3)
p <- ggplot(df, aes(SJR, mean.score)) +
  geom_point(alpha=0.3, size = 3) +
  stat_smooth(method = lm,fullrange=TRUE) +
  theme_classic()
ggExtra::ggMarginal(p, type = "histogram")
dev.off()

```

## Article citations

Get the number of citations per article.

```{r,cites}

if (file.exists("../data/cites_data.rds") ) {
  cites <- readRDS("../data/cites_data.rds")
} else {
  stop("citation data not found. Please run 'citation_fetching.Rmd' to obtain updated citation info")
}
cites$PMID <- gsub("^","PMID",cites$PMID)
xm <- merge(x,cites,by.x="Pubmed.Central.ID",by.y="PMCID")
dim(xm)
head(xm)

```

Plot the number of citations against analysis scores

```{r,citesplot}

mylm2 <- lm (xm$scores ~ xm$NumCites)
plot(xm$NumCites,xm$scores, col = rgb(red=0,blue=0,green=0,alpha=0.2) , pch=19,bty="n",
  xlab="Num cites",ylab="score",main="Association of analysis scores with number of citations")
abline(mylm2,col="red")

cor.test(xm$scores , xm$NumCites, method="pearson")

cor.test(xm$scores , xm$NumCites, method="spearman")

mylm2 <- lm (xm$scores ~ log2(xm$NumCites+1))
plot(log2(xm$NumCites),xm$scores, col = rgb(red=0,blue=0,green=0,alpha=0.2) , pch=19,bty="n",
  xlab="log2(citations)",ylab="score",main="Analysis scores vs. log2(citations)")
abline(mylm2,col="red")

png("images/cites1.png",width=350,height=300)
par(mar=c(5,5,3,1))
plot(log2(xm$NumCites),xm$scores,col = rgb(red=0,blue=0,green=0,alpha=0.2) , pch=19,bty="n",
  xlab="log2(citations)",ylab="score",main="Analysis scores vs. log2(citations)")
abline(mylm2,col="red")
dev.off()

cor.test(xm$scores , log2(xm$NumCites+1), method="pearson")

cor.test(xm$scores , log2(xm$NumCites+1), method="spearman")

pdf("images/cites1.pdf",width=3.5,height=3)
par(mar=c(5,5,3,1))
plot(log2(xm$NumCites),xm$scores,col = rgb(red=0,blue=0,green=0,alpha=0.2) , pch=19,bty="n",
  xlab="log2(citations)",ylab="score",main="Analysis scores vs. log2(citations)")
abline(mylm2,col="red")
dev.off()

df <- data.frame("log2.citations" = log2(xm$NumCites+1), "score" = xm$scores )

p <- ggplot(df, aes(log2.citations, score)) +
  geom_point(alpha=0.1, size = 3) +
  stat_smooth(method = lm,fullrange=TRUE) +
  theme_classic()
ggExtra::ggMarginal(p, type = "histogram")

pdf("images/cites2.pdf",width=4,height=3)
p <- ggplot(df, aes(log2.citations,score)) +
  geom_point(alpha=0.1, size = 3) +
  stat_smooth(method = lm,fullrange=TRUE) +
  theme_classic()
ggExtra::ggMarginal(p, type = "histogram")
dev.off()

```

## Session information

```{r,session}

sessionInfo()

```